\section{Training Plan Review}\label{sec:training}

\subsection{Training Provided}\label{sub:training:provided}

The training provided thus far has take two form --- technical training
undertaken as taught modules on the CDT programme and non-technical training
pursued via other avenues.
Aside from the core set of modules undertaken for the CDT, I have also
undertaken the following option modules:
\begin{itemize}
    \item Big Data Consumer Analytics
    \item Parallel and Concurrent Programming (COMP5811M)
    \item Programming for Geography Information Analysis: Advanced Skills
        (GEG5790M)
\end{itemize}
These have aimed to gain a better understanding data scientific tools and to
improve my programming knowledge.
The other non-academic forms have training that I have undertaken are as
follows:
\begin{itemize}
    \item Foundations in Teaching (University of Leeds) --- Provided with an
        introduction to teaching and demonstrating as a PGR.
    \item Identifying Impact Goals (University of Leeds) --- Learned about
        reframing research with a view to achieving impact.
    \item Networking (University of Leeds) --- Gain skills for professional
        networking at conferences.
    \item Time Management (University of Leeds) --- Reviewed methods for time
        management.
    \item Literature Searching (University of Leeds) --- Learned about how to
        undertake more thorough and systematic literature reviews.
    \item Facilitator training (Turing Institute Data Study Group) --- Gained
        interpersonal skills and a greater understanding of how to manage a
        research project.
\end{itemize}

\subsection{Training Required}\label{sub:training:required}

The training provided thus far has been substantial; consequently, comparatively
little time will be spent seeking training, with more time being allocated to
PhD research.
The training that shall be sought, however, will focus on writing:
\begin{itemize}
    \item Writing for Academic Publication --- The CDT is hosting a workshop on
        getting work published at its annual event this month which I plan to
        attend.
    \item Thesis Writing --- The ODPL offers a series of workshops on developing
        writing skills specifically focused on thesis writing.
\end{itemize}
